---
title: "Mapping flights with Leaflet and Shiny"
author: ""
date: ""
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Related to the last post, I was curious to see where one can fly from Brisbane. After a bit of research, I saw that data on [airports, airlines and routes](https://openflights.org/data.html) are available online. I also saw online some attempts 

Here is a map that uses shiny and leaflet in R to map all destinations from most airports in the world. The code, which is partly based on a [previous map made with ggplot](http://web.stanford.edu/~cengel/cgi-bin/anthrospace/great-circles-on-a-recentered-worldmap-in-ggplot), can be found below.

<iframe style="overflow:hidden" scrolling="yes" width="100%" height=600" frameBorder="0" src="https://accastonguay.shinyapps.io/flights_app/" > </iframe>



```{r eval = FALSE}
library("shiny")
library("leaflet")
library("dplyr")
library("geosphere")
library("plyr") 

airports <- read.table("https://raw.githubusercontent.com/jpatokal/openflights/master/data/airports.dat", 
                       header=F, sep=",")

flights <-  read.table("https://raw.githubusercontent.com/jpatokal/openflights/master/data/routes.dat", 
                       header=F,  sep=",")

airlines <-  read.table("https://raw.githubusercontent.com/jpatokal/openflights/master/data/airlines.dat", 
                        header=F, sep=",") %>% dplyr::select(V2, V4)

colnames(flights) <- c("Airline_code", "c2", "origin", "c4", "destination", "c6", "c7", "c8","c9")
all_data0 <- dplyr::select(flights, Airline_code, origin, destination) %>%
  dplyr::left_join(airports[, c("V3","V4", "V5", "V7","V8")],by = c("origin" = "V5")) %>%
  dplyr::left_join(airports[, c("V3","V4", "V5", "V7","V8")],by = c("destination" = "V5"))%>%
  dplyr::left_join(airlines, by = c("Airline_code"= "V4"))%>%
  mutate(city = paste(V3.x, ", ",     V4.x, sep = ""))


colnames(all_data0) <- c("Airline_code", "origin", "destination", "city1", "country",  "lat_ori", "lng_ori", "city_dest", "country_dest", "lat_dest", "lng_dest", "airline", "city")

all_data <- na.omit(all_data0)

all_data$city <- dplyr::recode(all_data$city,
                        "Östersund, Sweden" = "Oestersund, Sweden",
                        "Ängelholm, Sweden" = "Aengelholm, Sweden",
                        "Çorlu, Turkey" = "Corlu, Turkey")
                        
bootstrapPage(
  tags$style(type = "text/css", "html, body {width:100%;height:100%}"),
  leafletOutput("map", width = "100%", height = "100%"),
  absolutePanel(top = 60, right = 10,
                selectInput("origin","Origin",
                            selected = "Brisbane, Australia",
                            levels(factor(all_data$city))),
                # uiOutput("airline"),
                style = "opacity: 0.65; z-index: 1000;")
) 

function(input,output, session){
  
  Dataframe2 <- reactive({
    data <- all_data[all_data$city %in% input$origin,]})
  
  output$map <- renderLeaflet({
    data <- Dataframe2()
    
    df2<-gcIntermediate(as.matrix(data[,c("lng_ori", "lat_ori")]),
                        as.matrix(data[,c("lng_dest", "lat_dest")]),
                        n=100,
                        addStartEnd=TRUE,
                        sp=TRUE,
                        breakAtDateLine=FALSE)
    df2 <- as(df2, "SpatialLinesDataFrame")
    df2.ff <- fortify(df2)
    
    data$id <-as.character(c(1:nrow(data))) 
    gcircles <- merge(df2.ff, data, all.x=T, by="id")
    
    if (data$lng_ori[1] > 0){
      center <- data$lng_ori[1]
    }
    else{
      center <- data$lng_ori[1] + 360
    }
    
    gcircles$long.recenter <-  ifelse(gcircles$long  < center - 180 , gcircles$long + 360, gcircles$long)
    data$long.ori.recenter <-  ifelse(data$lng_ori  < center - 180 , data$lng_ori + 360, data$lng_ori)
    data$long.dest.recenter <-  ifelse(data$lng_dest  < center - 180 , data$lng_dest + 360, data$lng_dest)
    
    test_line <- sf::st_as_sf(gcircles, coords = c("long.recenter", "lat")) %>%
      dplyr::group_by(id, piece) %>%
      dplyr::summarize(do_union=FALSE) %>%
      sf::st_cast("LINESTRING") %>%
      dplyr::ungroup()
    
    test_line2 <- dplyr::left_join(test_line, data)
    
    labels <- sprintf(
      "<strong>%s, </strong> %s",
      data$city_dest, data$country_dest) %>% lapply(htmltools::HTML)
  
    m <- leaflet(data =  test_line2) %>%
      addTiles() %>%
      addProviderTiles(providers$CartoDB.Positron, group = "Carto DB Positron") %>%
      addProviderTiles(providers$CartoDB.DarkMatterNoLabels, group = "Carto DB dark")%>%
      addPolylines(weight = 1, opacity = 0.5, color = "#820a0a", label = ~airline) %>%
      addCircleMarkers(data = data, lng = ~long.ori.recenter, lat = ~lat_ori, radius =0.5, fillOpacity = 0.1,
                       weight = 2,opacity = 0.1, color = "red") %>%
      addCircleMarkers(data = data, lng = ~long.dest.recenter, lat = ~lat_dest, radius =1, label = labels) %>%
      addLayersControl(baseGroups = c("Carto DB Positron","Carto DB dark"))
    m
  })
}
```
